---
title: "Reproducible document"
author: "Stine Jorstad Bugge"
format: html
---

## Basics of R

```{r}
1 + 1
```

## Setup code chunk

```{r setup}
#| message: false
#| warning: false
library(tidyverse)
post_meal_data <- read_csv2(
  here::here("data/post-meal-insulin.csv")
)

small_post_meal_data <- select(post_meal_data, id, Age, BMI, Group, auc_pg, auc_ins, glykemi)

tidier_post_meal_data <- small_post_meal_data |>
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  ) |>
  filter(glykemi == 0) |>
  select(-glykemi)
tidied_post_meal_data <- tidier_post_meal_data |>
  mutate(
    insulin_glucose_ratio = auc_insulin / auc_glucose,
    Group = if_else(Group == "CTR", "Control", "First-degree relative")
  )
```
## Showing the data
```{r}
post_meal_data
```
```{r}
glimpse(post_meal_data)
```

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```
## Distribution of BMI

```{r}
#| fig-cap: "Distribution of BMI"
#| label: fig-bmi-histo
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

@fig-bmi-histo

## Exercise: discrete plots

```{r}
#| fig-cap: "Glykemi"
#| label: fig-glykemi-bar
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

See @fig-glykemi-bar above for a cool plot!

 linje 60 er en anden måde at sige linje 51 

```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = glykemi)) +
  geom_bar()
```

linje 69 fortæller R at glykemi-variablen skal være en karakter og ikke numerisk.

```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar()
```

```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```
Hvis du har brug for hjælp, så kan du skrive ?og den funktion du vil vide om i Console

## About me
- Stine Jorstad Bugge
- Steno Diabetes Center Odense
- Farmakologi, Odense Universitetshospital
- Syddansk Universitet

As a doctor I care for patients and want to help them navigate their disease, treatment and care. Therefore I work with an *intervention* aiming to bring doctors together to improve coordination and care for patients with **multimorbidity and diabetes**.

## Simple code
```{r}
3 * 3
```
## Testing for Git
testing testing one two three

## Putting two plots side by side 

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

```{r}
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
``` 
```{r side-by-side-figs}
#| label: fig-bmi-glykemi
#| fig-cap: "BMI and glykemi, side by side"
#| fig-subcap:
#| - "Distribution of BMI"
#| - "Number of those with glykemi"
#| layout-ncol: 2

ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()

ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

## Plotting two continous variables

```{r}
#| fig-cap: "Scatterplot showing BMI against AUC of blood glucose"
#| label: fig-bmi-auc-pg
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```

## Pushing changes to remoter

## Pulling changes locally

## Selecting data

```{r}
select(post_meal_data, Age)

select(post_meal_data, Age, BMI, Weight)

select(post_meal_data, -Group)

select(post_meal_data, starts_with("PG"))

select(post_meal_data, ends_with("DXA"))

select(post_meal_data, contains("0"))
```

## Renaming specific columns

```{r}
small_post_meal_data |>
  rename(auc_glucose = auc_pg)

small_post_meal_data |>
  rename(auc_glucose = auc_pg, auc_insulin = auc_ins)
```

## Exercise 11.6

## Select and rename tasks

```{r}
# Task 1.
post_meal_data |>
  select(contains("OGTT"))

# Task 2.
post_meal_data |>
  select(contains(".."), -contains("...")) |>
  rename(P.Glucose.minus5.OGTT = P.Glucose..5.OGTT) |>
  rename(Insulin.minus5.OGTT.X = Insulin..5.OGTT.X)

# Task 3.
basic_info <- select(post_meal_data, id, BMI, Length, Weight, Age)
rename(basic_info, Height = Length)
```

# Task 1.
post_meal_data |>
  select(contains("OGTT"))

# Task 2.
post_meal_data |>
  select(contains(".."), -contains("...")) |>
  rename(
    P.Glucose.minus5.OGTT = P.Glucose..5.OGTT,
    Insulin.minus5.OGTT.X = Insulin..5.OGTT.X
  )

# Task 3.
post_meal_data |>
  select(id, BMI, Length, Weight, Age) |>
  rename(Height = Length)

## Filtering data by row
```{r}
small_post_meal_data |>
  filter(glykemi == 1)

small_post_meal_data |>
  filter(BMI >= 25)

small_post_meal_data |>
  filter(Group == "CTR")
```

```{r}
TRUE & TRUE
```
```{r}
small_post_meal_data |>
  filter(Group == "CTR", BMI >= 25)

small_post_meal_data |>
  filter(Group == "CTR" | BMI >= 25)
```

## Modify column

```{r}
tidier_post_meal_data |>
  mutate(insulin_glucose_ratio = auc_insulin / auc_glucose, Group = if_else(Group == "CTR", "Control", "First -degree relative"))
```

## Practice with filter and mutate

```{r}
post_meal_data |>
  filter(BMI >= 20 & BMI <= 40 & Group == "FDR")
```

```{r}
post_meal_data |>
  mutate(
    percent_body_fat = (Fat.mass.DXA / Weight) * 100
  ) |>
  filter(percent_body_fat > 30)
```

## Visualising wth jitter plot
```{r}
tidied_post_meal_data |>
  mutate(overweight = if_else(BMI >= 25, "Overweight", "Not overweight")) |>
  ggplot(aes(x = overweight, y = insulin_glucose_ratio)) +
  geom_jitter()
```
## Plotting three or more variables

```{r}
tidied_post_meal_data |>
  filter(BMI < 30) |>
  ggplot(aes(x = Age, y = auc_insulin, colour = Group)) +
  geom_point()
```
```{r}
tidied_post_meal_data |>
  filter(BMI < 30) |>
  mutate(young = if_else(Age < 40, "Young", "Old")) |>
  ggplot(aes(x = BMI, y = insulin_glucose_ratio, colour = Group)) +
  geom_point() +
  facet_grid(rows = vars(young)) +
  labs(
    x = "Body mass index (BMI)",
    y = "Insulin to glucose ration"
  )
```
## Make multi-dimensional plots!

```{r}
#| label: fig-multidimensional
#| fig-cap: "Two, mult-dimensional plots of our data."
#| fig-subcap:
#|   - "Distribution of percent body fat by group and age category."
#|   - "Relationship between BMI and bone mineral density, split by insulin sensitivity."
#| layout-ncol: 2
# Task 1.
post_meal_data |>
  mutate(
    percent_body_fat = (Fat.mass.DXA / Weight) * 100,
    young = if_else(Age < 40, "Young", "Old")
  ) |>
  ggplot(aes(x = Group, y = percent_body_fat, colour = young)) +
  geom_jitter(width = 0.2)
```

```{r}
post_meal_data |>
  mutate(
    insulin_glucose_ratio = (auc_ins / auc_pg),
    insulin_insensitive = if_else(insulin_glucose_ratio > mean(insulin_glucose_ratio), "Insensitive", "Sensitive")
  ) |>
  ggplot(aes(x = BMI, y = Bone.mineral.DXA)) +
  geom_point() +
  geom_smooth() +
  facet_grid(cols = vars(insulin_insensitive))
```
## Changing colours

```{r}
post_meal_data |>
  mutate(hyperglycemia = if_else(glykemi == 1, "Yes", "No")) |>
  ggplot(aes(x = Group, fill = hyperglycemia)) +
  geom_bar(position = position_dodge()) +
  scale_fill_viridis_d(option = "cividis")
```
```{r}
post_meal_data |>
  ggplot(aes(x = auc_pg, y = auc_ins, colour = BMI)) +
  geom_point() +
  scale_colour_viridis_c()
```
```{r}
post_meal_data |>
  ggplot(aes(x = auc_pg, y = auc_ins, colour = BMI)) +
  geom_point() +
  scale_colour_viridis_c(option = "inferno")
```

## Calculating summary statistics
# Split-apply-combine

```{r}
tidied_post_meal_data |>
  summarise(
    max_bmi = max(BMI),
    min__bmi = min(BMI)
  )
```

## Summarise with splitting

```{r}
tidied_post_meal_data |>
  group_by(Group) |>
  summarise(
    mean_age = mean(Age),
    mean_bmi = mean(BMI)
  ) |>
  ungroup()
```
