---
title: "Reproducible document"
author: "Stine Jorstad Bugge"
format: html
---

## Basics of R

```{r}
1 + 1
```

## Setup code chunk

```{r setup}
#| message: false
#| warning: false
library(tidyverse)
post_meal_data <- read_csv2(
  here::here("data/post-meal-insulin.csv")
)
```
## Showing the data
```{r}
post_meal_data
```
```{r}
glimpse(post_meal_data)
```

```{r}
ggplot(post_meal_data, aes(x = BMI)) + 
  geom_histogram()
```
## Distribution of BMI

```{r}
#| fig-cap: "Distribution of BMI"
#| label: fig-bmi-histo
ggplot(post_meal_data, aes(x = BMI)) + 
  geom_histogram()
```

@fig-bmi-histo

## Exercise: discrete plots

```{r}
#| fig-cap: "Glykemi"
#| label: fig-glykemi-bar
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

See @fig-glykemi-bar above for a cool plot!


## About me
- Stine Jorstad Bugge
- Steno Diabetes Center Odense
- Farmakologi, Odense Universitetshospital
- Syddansk Universitet

As a doctor I care for patients and want to help them navigate their disease, treatment and care. Therefore I work with an *intervention* aiming to bring doctors together to improve coordination and care for patients with **multimorbidity and diabetes**.

## Simple code
```{r}
3 * 3
```
## Testing for Git
testing testing one two three

