---
title: "Reproducible document"
author: "Stine Jorstad Bugge"
format: html
---

## Basics of R

```{r}
1 + 1
```

## Setup code chunk

```{r setup}
#| message: false
#| warning: false
library(tidyverse)
post_meal_data <- read_csv2(
  here::here("data/post-meal-insulin.csv")
)
```
## Showing the data
```{r}
post_meal_data
```
```{r}
glimpse(post_meal_data)
```

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```
## Distribution of BMI

```{r}
#| fig-cap: "Distribution of BMI"
#| label: fig-bmi-histo
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

@fig-bmi-histo

## Exercise: discrete plots

```{r}
#| fig-cap: "Glykemi"
#| label: fig-glykemi-bar
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

See @fig-glykemi-bar above for a cool plot!

 linje 60 er en anden måde at sige linje 51 

```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = glykemi)) +
  geom_bar()
```

linje 69 fortæller R at glykemi-variablen skal være en karakter og ikke numerisk.

```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar()
```

```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```
Hvis du har brug for hjælp, så kan du skrive ?og den funktion du vil vide om i Console

## About me
- Stine Jorstad Bugge
- Steno Diabetes Center Odense
- Farmakologi, Odense Universitetshospital
- Syddansk Universitet

As a doctor I care for patients and want to help them navigate their disease, treatment and care. Therefore I work with an *intervention* aiming to bring doctors together to improve coordination and care for patients with **multimorbidity and diabetes**.

## Simple code
```{r}
3 * 3
```
## Testing for Git
testing testing one two three

## Putting two plots side by side 

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

```{r}
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
``` 
```{r side-by-side-figs}
#| label: fig-bmi-glykemi
#| fig-cap: "BMI and glykemi, side by side"
#| fig-subcap:
#| - "Distribution of BMI"
#| - "Number of those with glykemi"
#| layout-ncol: 2

ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()

ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

## Plotting two continous variables

```{r}
#| fig-cap: "Scatterplot showing BMI against AUC of blood glucose"
#| label: fig-bmi-auc-pg
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) + 
  geom_point() + 
  geom_smooth()
```

## Pushing changes to remote

## Pulling changes locally
